\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-24 coqimpl.tex
Reply part of the comments.

Despite possible difficulties for future maintenance, this approach has
various advantages: 1. it is the easiest and the most accessible way to
prototype as it relieves us the necessity of familarity of Coq base,
especially for the implementation of the module and functor; 2. we have
a clear definition of trusted-base---the whole Coq; 3. it is easy to
debug---we just need to check the translated commands; 4. it can be
well-incorporated with the existent tools like VSCoq; 5. it is more
accessible for the interested audience who can then easily adapt our
plugin and give a try---otherwise they have to download and re-build the
whole customized Coq; 6. it is more stable because the surface syntax of
Coq should be stable across different versions; 7. still, this plugin
can capture the key ingredients of implementing family polymorphism
inside Coq and act as a reference for guiding an appropriate
implementation of Family Polymorphism inside all sorts of proof
assistants.\YZ{Does the Coq implementation perform type checking to
prevent illegal uses fam poly?}\EDJreply{Yes. Because I define feature of fampoly as a direct translation into Coq feature.  The fampoly is like a shorthand into multiple Coq commands. \\
Do you find translation from my current metatheory into MLTT (without linkage) helpful? \\
Actually I am also not sure what this question is aiming towards. \\
Are you complaining there is no formal specification of my fampoly feature in paper and thus there is no proof my Coq surface syntax is related to metatheory? If so, I thought our Coq plugin is only advertising how powerful Fampoly is to solve expression problem in mechanized proving. I can decouple the relationship between our plugin and our metatheory if you find the gap is really there... And I can only claim metatheory is about a partial result on incorporating fampoly with dependent type, Coq plugin is using two example to illustrate how mechanized proving can get benefit and our Coq plugin is only inspired by our metatheory (and not claim any relationship). Doing so would require rearrange the metatheory section before this section. \\
But I don't think the gap is there at all because if you want all possible gaps to disappear, then the action of using debruijn is wrong because we need to prove the one using debruijn and the surface syntax that is not using debruijn are ``equivalent''. What's more, module is not sigma type because Module is not a term in Coq and thus there is no formalism of Module at all. Let alone I don't see the plausibility of mechanizing dependent type without debruijn. \\ 
Please comment back so that I know what corresponding modification is needed on the paper. 
}\YZreply{My original question was more directed towards the fact
that currently type checking is deferred to the Coq base. Generally,
when one uses a statically typed language as a compilation target, they
also want to have type checking at the source level, (1) because there
could be type errors at the source level that could not otherwise be
caught at the target level and (2) because it allows better error messages.
So it needs to be addressed why deferring type checking to Coq is OK.}

\textbf{Family compiled into module.} This is the main idea of compiling
family components into Coq's primitives: we compile families into Coq
modules, family types into module types, and the context of a given term
into parameters of the module. Unlike module, we need to faithfully
reflect the \textit{late binding} nature of families: for each defining
field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
choose module and module parameters to achieve this wrapping---each
family type in the context of the judgement will be one module parameter
with the compiled module type, and thus each field of the family will
compile into a functor parametrized by its context. Doing so we can also
get rapid feedback from type checking when the users are defining each
field interactively with Coq.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}[!htb]
  \begin{minipage}[t]{0.25\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family A.
Field a : nat 
  := 1.
Field b 
  : self_A.a = self_A.a
  := eq_refl.
Final Field a' : nat 
  := 1. 
Field b'
  : self_A.a' = 1
  := eq_refl.
EndFamily.
\end{minted}
  \end{minipage}\YZ{Can the self's be omitted?}\EDJreply{Yes. But in our current plugin implementation it is not omitted. I think altering example is fine. If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}
  \begin{minipage}[t]{0.4\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_5 (self_A: EmptySig).
Parameter (a : nat). End a_5.
Module Type a_6 := a_5 EmptyMod.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = self_A.a 
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: when the user
interactively inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_5, a_6"
three components.
{The "EmptySig" is just empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_10"\YZ{should this be a_4?} is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking and overriding as in \ref{langdesign-req0}---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. This compiled module type serves as an
abstraction that enables overriding.

Now that we know a family data structure is actually a list of compiled
modules, we may want to aggregate them into one complete compiled
module. We achieve that by repeatedly using Coq's "Include" command as
the example illustrated. Note that, in our example, we include two
functors without arguments---this is due to a perculiarity of Coq's
"Include": non-instantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.\YZ{'For example, ...'}\EDJreply{I don't think explaining a Coq feature inside our paper is a good idea? But I have added this in the following line, please check. Keep it and resolve the comment if you find this ok. }
For example, "b_7" requires a module of type "a_6". Inside module "A", when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat} inside the current surrounding "A". Then when we \mintinline{Coq}{Include b_7} without specifying the module parameters, Coq will try to find fields in the surrounding module, (i.e. module "A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by the \mintinline{Coq}{a : nat} that is included earlier.
The resulting module is just like mundane Coq's module and thus
enjoy all the computational mechanism from Coq for free.

@@

To fix that, we need more information on "a" than its type---we
need to expose the concrete definition "a = 1" to the context of "b".
\YZ{I had a hard time working through these two paras. I guess they are
saying that there is a need to expose definitional equalities, and
therefore we need non-overridables. This causal flow should be made
more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph.}

>>>>

Despite possible difficulties for future maintenance, this approach has
various advantages: 1. it is the easiest and the most accessible way to
prototype as it relieves us the necessity of familarity of Coq base,
especially for the implementation of the module and functor; 2. we have
a clear definition of trusted-base---the whole Coq; 3. it is easy to
debug---we just need to check the translated commands; 4. it can be
well-incorporated with the existent tools like VSCoq; 5. it is more
accessible for the interested audience who can then easily adapt our
plugin and give a try---otherwise they have to download and re-build the
whole customized Coq; 6. it is more stable because the surface syntax of
Coq should be stable across different versions; 7. still, this plugin
can capture the key ingredients of implementing family polymorphism
inside Coq and act as a reference for guiding an appropriate
implementation of Family Polymorphism inside all sorts of proof
assistants.\YZ{Does the Coq implementation perform type checking to
prevent illegal uses fam poly?}\EDJreply{Yes. Because I define feature of fampoly as a direct translation into Coq feature.  The fampoly is like a shorthand into multiple Coq commands. \\
Do you find translation from my current metatheory into MLTT (without linkage) helpful? \\
Actually I am also not sure what this question is aiming towards. \\
Are you complaining there is no formal specification of my fampoly feature in paper and thus there is no proof my Coq surface syntax is related to metatheory? If so, I thought our Coq plugin is only advertising how powerful Fampoly is to solve expression problem in mechanized proving. I can decouple the relationship between our plugin and our metatheory if you find the gap is really there... And I can only claim metatheory is about a partial result on incorporating fampoly with dependent type, Coq plugin is using two example to illustrate how mechanized proving can get benefit and our Coq plugin is only inspired by our metatheory (and not claim any relationship). Doing so would require rearrange the metatheory section before this section. \\
But I don't think the gap is there at all because if you want all possible gaps to disappear, then the action of using debruijn is wrong because we need to prove the one using debruijn and the surface syntax that is not using debruijn are ``equivalent''. What's more, module is not sigma type because Module is not a term in Coq and thus there is no formalism of Module at all. Let alone I don't see the plausibility of mechanizing dependent type without debruijn. \\ 
Please comment back so that I know what corresponding modification is needed on the paper. 
}\YZreply{My original question was more directed towards the fact
that currently type checking is deferred to the Coq base. Generally,
when one uses a statically typed language as a compilation target, they
also want to have type checking at the source level, (1) because there
could be type errors at the source level that could not otherwise be
caught at the target level and (2) because it allows better error messages.
So it needs to be addressed why deferring type checking to Coq is OK.}\EDJreply{Good point. I think (1) is largely mediated by the fact the semantic of family is very close to that of the module. And the supporting feature is simple enough that there won't appear some problems like type error at source-level but not target level(2) is happening for sure but some basic error message is almost "transparent" i.e. the error-messages from Coq can be directly interpreted for the user. But my justification is solely empirical and as you know the empirical study for this paper is not really strong enough to support anything.}

\textbf{Family compiled into module.} This is the main idea of compiling
family components into Coq's primitives: we compile families into Coq
modules, family types into module types, and the context of a given term
into parameters of the module. Unlike module, we need to faithfully
reflect the \textit{late binding} nature of families: for each defining
field $ .. \cL\sigma \vdash t : T $, we need to compile $t$ using
``universal-quantifier-wrapped term'' : $.. \vdash \lambda t : \forall
\cL \sigma. T$. However, instead of using Coq's universal quantifier, we
choose module and module parameters to achieve this wrapping---each
family type in the context of the judgement will be one module parameter
with the compiled module type, and thus each field of the family will
compile into a functor parametrized by its context. Doing so we can also
get rapid feedback from type checking when the users are defining each
field interactively with Coq.
% Insert one pseudo-code example for explanation of the mechanism
\begin{figure}[!htb]
  \begin{minipage}[t]{0.25\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Family A.
Field a : nat 
  := 1.
Field b 
  : a = a
  := eq_refl.
Final Field a' : nat 
  := 1. 
Field b'
  : a' = 1
  := eq_refl.
EndFamily.
\end{minted}
  \end{minipage}\YZ{Can the self's be omitted?}\EDJreply{Yes. But in our current plugin implementation it is not omitted. I think altering example is fine. If you think some small distance between example and the current implementation is fine, please comment back and I will remove all the self_}\YZreply{If omitting the self's poses no technical challenges, then I'd say it's fine to also omit them in this figure.}\EDJreply{Omitting self_ is easy and not much engineering effort is required. But omitting self_ will lead to `A.a' instead of `a' actually. So if we really want `a' as you last time mentioned, then a resolution is required and more engineering effort is required. But I don't see any technical difficulties. I have removed all the self, please check.}
  \begin{minipage}[t]{0.4\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module a_4 (self_A: EmptySig).
Definition a : nat := 1. End a_4.
Module Type a_5 (self_A: EmptySig).
Parameter (a : nat). End a_5.
Module Type a_6 := a_5 EmptyMod.
\end{minted}
  \end{minipage}
  \begin{minipage}[t]{0.3\linewidth}
\begin{minted}[fontsize=\footnotesize]{Coq}
Module b_7 (self_A: a_6).
Definition b 
  : self_A.a = self_A.a 
  := eq_refl. End b_7.

@@

% Concrete details of compilation of a family term data structure
Taking \cref{fig:plugin-example1} as an example: when the user
interactively inputs the line "Field a : nat := ..", our plugin will
translate this statement into the "Module a_4", "Module Type a_5, a_6"
three components.
{The "EmptySig" is just empty module type.} Then,
these modules and module types will be generated and type-checked by Coq
immediately. The compiled "Module a_4"\YZ{should this be a_4?}\EDJreply{Yes. Thanks!} is then part of the context of
"Field b" (as the module type of "self_A") for "b" to refer to. The
internal representation of family "A" is the list of these
compiled modules and module types. We can achieve incremental
checking and overriding as in \ref{langdesign-req0}---for example, if we
want to override "Field a", we simply replace that compiled module with
the new module inside the list. This compiled module type serves as an
abstraction that enables overriding.

Now that we know a family data structure is actually a list of compiled
modules, we may want to aggregate them into one complete compiled
module. We achieve that by repeatedly using Coq's "Include" command as
the example illustrated. Note that, in our example, we include two
functors without arguments---this is due to a perculiarity of Coq's
"Include": non-instantiated parameters will be automatically
instantiated by the appropriate fields in the surrounding defining
module.\YZ{'For example, ...'}\EDJreply{I don't think explaining a Coq feature inside our paper is a good idea? But I have added this in the following line, please check. Keep it and resolve the comment if you find this ok. }
For example, "b_7" requires a module of type "a_6". Inside module "A", when "Include a_4" is done, we will have a field \mintinline{Coq}{a : nat} inside the current surrounding "A". Then when we \mintinline{Coq}{Include b_7} without specifying the module parameters, Coq will try to find fields in the surrounding module, (i.e. module "A") to instantiate \mintinline{Coq}{self_A: a_6}. This is satisfied by the \mintinline{Coq}{a : nat} that is included earlier.
The resulting compiled module is just like mundane Coq's module and thus maintains all the computational mechanism from Coq, fulfilling \ref{langdesign-req0b}.

@@

To fix that, we need more information on "a" than its type---we
need to expose the concrete definition "a = 1" to the context of "b".
\YZ{I had a hard time working through these two paras. I guess they are
saying that there is a need to expose definitional equalities, and
therefore we need non-overridables. This causal flow should be made
more explicit.}\EDJreply{Good suggestion. I try to add this casual flow into the first line of the paragraph. Please check and resolve if find satisfied.}

### 2022-06-24 lang-design.tex
Reply part of the comments.

When adapting Family Polymorphism into dependent types, we choose to
focus only on the essence of family (and inheritance)
structure in \citet{zm2017}, and thus a lot of unrelated features, like
Interface, will be removed. In this case, it will look like module with
late-binding and inheritance as code and proof reuse for module. We also
need it to have good compatibility with inductive types, because we
don't want to lose too much reasoning power of Coq\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}
(and its tactic programming).

@@

We use "Family" and "EndFamily" to scope.
Each field in a family is declared and defined as a "Field".\YZ{This sentence sounds a bit recursive}
We can define nested families.
Family will also be used to encode recursive handlers. Each case
handling of one inductive type will be aggregated into one family, then
"FRecursor" will be used to construct a recursor using these handlers.
For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
"handlers" will generate a new field "recur" that is doing induction
(using elimination principle of "T") using motive "f", and the handlers
are located in the family "handlers".

>>>>

When adapting Family Polymorphism into dependent types, we choose to
focus only on the essence of family (and inheritance)
structure in \citet{zm2017}, and thus a lot of unrelated features, like
Interface, will be removed. In this case, it will look like module with
late-binding. Inheritance can be considered as code and proof reuse mechanism for module. We also
need it to have good compatibility with inductive types, because we
don't want to diverge too much from the routine Coq programming experience\YZ{By 'reasoning power of Coq', I suppose you mean the expressive power of Gallina?}\EDJreply{I think the sentence last time was not really making sense because using Church Encoding of inductive doesn't sabotage the expressiveness. I rephrase it now, please check.}
(and its tactic programming).

@@

We use "Family" and "EndFamily" to scope, and "Field" is used to declare new field members in family.\YZ{This sentence sounds a bit recursive}\EDJreply{What about now?}
We can define nested families.
Family will also be used to encode recursive handlers. Each case
handling of one inductive type will be aggregated into one family, then
"FRecursor" will be used to construct a recursor using these handlers.
For example, \mintinline{Coq}{FRecursor recur about T motive f} "using"
"handlers" will generate a new field "recur" that is doing induction
(using elimination principle of "T") using motive "f", and the handlers
are located in the family "handlers".

### 2022-06-24 syntax.tex
Syntactical Translation added. Please check.
Reply part of the comments.

Now we show selected typing rules: we will have dependent function types
and dependent pair types. Dependent pair type is used to model Coq
module, which is useful when we need to model the ``compilation'' from
linkages to modules. We also have $\beta,\eta$-rule for both dependent
function type and dependent pair type.

@@

\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{Working on it, don't resolve}

>>>>

Now we show selected typing rules: we will have dependent function types
and dependent pair types. Dependent pair type is used to model Coq
module, which is useful when we need to model the ``compilation'' from
linkages to modules. We also have $\beta,\eta$-rule for both dependent
function type and dependent pair type. The subsitution law for these two require \textit{de bruijn lifting}: for $\goodTerm{\Gamma}{\sigma}{\Delta}$, we define its de bruijn lifting $\goodTerm{\Gamma,A[\sigma]}{\sigma^\uparrow}{\Delta, A}$ given $\goodType{\Delta}{A}{}$.

@@

\YZ{Can we have a more concrete example of how the judgment is useful?}\EDJreply{it is done in the example section, please check.}
